Synthetic tabular data is often needed in settings where realistic examples are
useful for tests, demonstrations, notebooks, and exploratory modelling, but
direct reuse of source records is constrained by governance, licensing, privacy,
or operational policy. We present DataFrameSampler, a lightweight Python
library for mixed-type pandas dataframes that addresses this practical use case
without claiming formal privacy protection or general-purpose generative
optimality. The method constructs a fully numerical latent representation by
fitting one supervised Neighborhood Components Analysis block for each
categorical column, treating binary variables as categorical even when they are
stored numerically, and always excluding the target column from the context used
to predict it. Generation then applies a local mutual-neighbour translation in
the learned latent space: each generated row is constructed as an anchor row
modified by an observed displacement between two neighbouring rows. The sampler
rejects and retries numerically implausible candidates under fitted constraints
and decodes the result with inverse numeric scaling and probabilistic
random-forest categorical decoders. More
precisely, DataFrameSampler is intended to make a small, inspectable
construction feasible for low-stakes example-data workflows; therefore we
structure the paper as a claim-specific evidence argument rather than as a
single aggregate benchmark. The diagnostics include NCA-block ablations,
decoder probability checks, frozen-Isomap insertion stress, and a limited CTGAN
reference comparison. They support only bounded empirical claims: the method
has no global objective, no formal manifold-preservation guarantee, and no
privacy guarantee.
