\section{Evaluation Framework}
\label{sec:evaluation-framework}

The evaluation is organized around claims rather than around a single
leaderboard. This choice is central to the paper because the intended
contribution is bounded: DataFrameSampler is proposed as an inspectable
example-data generator, not as a general synthetic-data optimum, a privacy
mechanism, or a guaranteed manifold sampler. The primary claim is practical
use: the method should be useful when low setup, inspectability, and moderate
runtime matter. The secondary claim is mechanistic: the supervised categorical
latent blocks should improve the local structure needed by the displacement
operator. Consequently, each empirical quantity is assigned a role and a
specific failure mode.

\begin{table}[t]
\centering
\caption{Claim-specific evidence plan. Each row names the failure mode that the
evidence is meant to prevent, the role of the evidence in the paper's argument,
and the conclusion that remains allowable. Takeaway: the paper is structured as
an argument for practical, inspectable example generation, not as a general
synthetic-data optimality claim.}
\label{tab:claim-evidence}
\small
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{@{}llll@{}}
\toprule
Claim & Failure prevented & Evidence role & Instrument \\
\midrule
Practical usability & Hidden tuning or hidden cost & Primary & API, CLI, runtime, setup sensitivity \\
Data fidelity & Copying, separability, broken marginals & Primary & Neighbour, discrimination, distribution tests \\
Downstream utility & Plausible rows that harm tasks & Primary & Utility lift on held-out real tasks \\
NCA mechanism & Behaviour due to unrelated components & Supporting & NCA-vs-PCA, majority, raw-context checks \\
Decoder probabilities & Unmeasured probability quality & Supporting for reconstruction; boundary for calibrated draws & Log loss, Brier, calibration error \\
Analogy transport & Heuristic mistaken for guarantee & Boundary & Bootstrap, interpolation, hull, Isomap stress \\
Class rebalancing & Oversampling adjacency left untested & Boundary & Fixed minority augmentation diagnostic \\
Privacy and constraints & Diagnostics mistaken for release safety & Boundary & Row proximity, overlap, range checks \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

This framework also determines the allowable conclusion. If primary evidence
is missing, the practical claim must be weakened. If supporting evidence is
missing, the mechanism remains plausible rather than supported. If boundary
evidence is missing, the scope of use should be narrowed. The experiments
therefore ask whether the sampler behaves usefully under several diagnostic
views at once, not whether one favourable metric can be selected after the
fact. The current evidence supports limited setup and setup sensitivity, but
not a full equal-budget hyperparameter optimization claim against all
alternatives; that stronger claim remains outside the paper's evidence.
