\section{Conclusion}

DataFrameSampler provides a lightweight mixed-type dataframe sampler based
on iterative supervised NCA latent blocks and local mutual-neighbour transport.
The resulting estimator has a direct sklearn-style API and a file-oriented CLI,
but its modelling claim is intentionally narrow: it is a simple and inspectable
tool for example-data workflows, not a formally private or globally optimised
synthetic-data mechanism. Its most direct interpretability interface is the
analogy trace: each generated row is an anchor modified by an observed
neighbour-pair displacement in the fitted latent space.

The claim-specific evidence structure makes this conclusion more precise.
Categorical NCA blocks are supported only where they improve held-out
categorical prediction against simple baselines, decoder probabilities should
still be interpreted through their measured diagnostic quality, and manifold
evidence should be read as an insertion-stress diagnostic rather than as a
preservation guarantee. The Adult CTGAN reference reinforces the same
positioning: modern high-capacity tabular generators are relevant reference
points, but the paper's contribution is not to win a leaderboard against them.
It is to occupy an inspectable, low-setup point in the design space and to make
its diagnostics explicit. Future work should therefore focus on stability of the
iterative latent construction, stronger constraint-aware decoding,
high-cardinality preprocessing, and alternatives to independent per-column
decoders.
