\section{Discussion}

DataFrameSampler is designed for transparency and ease of use rather than for
formal generative modelling optimality. This is a deliberate trade-off. In many
business and medical software contexts, the most useful synthetic data tool is
not necessarily the most expressive generator, but the one that teams can run,
inspect, explain, and constrain. A local-neighbour bin-space mechanism gives
users a concrete account of how examples are created, which can make review and
failure analysis more realistic than with opaque generated rows.

This transparency should not be confused with privacy. The bin-level decoder
can emit exact source values for columns that are not replaced, which makes the
generated output unsuitable as a privacy mechanism by itself. The
sensitive-value replacement layer addresses one narrow failure mode: exact
re-emission of selected source values. It should be understood as a practical
safeguard for direct identifiers, not as a comprehensive anonymization method.
Business and medical deployments should still apply their normal governance,
data protection, and disclosure review procedures.

The method also depends on the quality of the latent representation. If
categorical values are represented only by frequency, semantically unrelated
categories with similar frequencies can become indistinguishable. Helper-column
embeddings address this limitation when meaningful numeric context exists, but
they also require appropriate configuration. The optional LLM-assisted mode is
intended to reduce configuration burden, not to replace user review. Its
recommendations are most useful when column names and example values reveal the
role of each field; they may be poor when names are opaque, columns are heavily
coded, or domain constraints are absent from the dataframe profile.

Approximate nearest-neighbour backends improve scalability, but can alter the
local graph used by the sampler. For this reason, exact and scikit-learn
backends remain useful baselines, especially for small datasets and experiments
where reproducibility is more important than throughput.

The preliminary Adult and Titanic comparisons in
Figures~\ref{fig:baseline-similarity} and
\ref{fig:configuration-competitors}, together with the distribution dashboard
in Figure~\ref{fig:distribution-dashboard} and the cost--similarity summary in
Figure~\ref{fig:utility-cost-frontier}, illustrate why the paper should avoid
a single leaderboard claim. Row bootstrap is strong on several similarity
metrics because it directly reuses source rows, but that behaviour is precisely
what limits it as a synthetic example generator for governed workflows.
Independent column sampling can preserve univariate distributions while
weakening association structure. DataFrameSampler configurations introduce a
different trade-off: they are slower than the simplest baselines, but the
generation mechanism remains inspectable and configuration choices visibly
change the measured behaviour.

The intended use cases are therefore pragmatic: test fixtures, workflow
demonstrations, training material, integration tests, dashboard prototypes, and
early-stage modelling experiments. The method should not be presented as a
drop-in replacement for a validated clinical data simulator, a regulated
statistical disclosure control process, or a differentially private synthetic
data release.

Future work includes more systematic evaluation, stronger privacy-aware
decoding strategies, constraint-aware sampling, richer LLM prompts that capture
domain constraints, and stronger model serialization formats. The current
fitted model is serialized with pickle, which should only be loaded from
trusted sources.
