\section{Planned Experiments}

This section is a stub for the empirical evaluation. The planned evaluation
will compare generated and source data along three axes.

\subsection{Distributional Similarity}

For numeric columns, planned metrics include univariate summary statistics,
histogram overlap, Wasserstein distance, and pairwise correlation differences.
For categorical columns, planned metrics include category coverage, frequency
divergence, and rare-category retention.

\subsection{Downstream Predictive Utility}

For supervised datasets, we plan to train predictive models on generated data
and evaluate them on held-out real data. This train-on-synthetic, test-on-real
protocol will provide a practical measure of whether generated samples preserve
useful predictive structure.

\subsection{Ablations}

Planned ablations include:

\begin{itemize}
  \item frequency encoding versus helper-column embedding for categorical
  columns;
  \item PCA, MDS, and nonlinear embedding choices;
  \item exact versus approximate nearest-neighbour backends;
  \item mutual-neighbour sampling with and without fallback edges.
\end{itemize}

\subsection{Datasets}

Candidate datasets include small public mixed-type tabular datasets and the
example datasets distributed with the project notebook. Any reported results
will clearly distinguish demonstration datasets from benchmark datasets.

\begin{table}[t]
\centering
\caption{Planned evaluation summary. Values are placeholders.}
\label{tab:planned-evaluation}
\begin{tabular}{lll}
\toprule
Dataset & Task & Metrics \\
\midrule
Heart & Classification & Accuracy, distribution shift \\
Billionaires & Mixed-type generation & Marginals, correlations \\
Synthetic correlated & Sanity check & Correlation retention \\
\bottomrule
\end{tabular}
\end{table}
