\section{Failure Modes and Limitations}
\label{sec:failure-modes}

The method has several predictable failure modes, and these boundaries are part
of the contribution. Because the sampler has no global likelihood, no joint
constraint solver, and no privacy mechanism, its appropriate use depends on
whether those limitations are acceptable for the workflow.

\paragraph{High-cardinality categories.}
Identifiers and near-identifiers can create large sparse one-hot blocks and
categorical targets that are hard to decode meaningfully. The implementation
warns on high cardinality, but it does not infer which columns are identifiers.
Such columns should be removed, grouped, or otherwise handled by the user when
exact source-value reuse is unacceptable.

\paragraph{Rare categories.}
Rare labels can be underrepresented by the learned decoder, especially when
they appear too infrequently to define stable neighbourhoods or calibrated
probabilities. Category coverage and rare-category preservation are therefore
reported as boundary diagnostics rather than assumed from the model form.

\paragraph{Cross-column constraints.}
The categorical decoders are trained independently, conditional on the final
latent context with the target block removed. It can therefore occur that each
decoded column is locally plausible while the combined row violates a
multi-column domain constraint. The fitted range and numeric standard-deviation
checks reduce some numeric failures, but they are not a general constraint
language.

\paragraph{Decoder uncertainty.}
Random-forest probabilities are empirical model outputs. Optional calibration
can improve probability interpretation when enough data are available, but
calibration itself must be measured. Poor log loss, Brier score, or expected
calibration error should be read as a warning about sampled categorical values,
not as a contradiction of the practical-use claim.

\paragraph{Transport instability.}
The local displacement rule assumes that the fitted latent geometry is locally
regular. If NCA blocks are unstable, if nearest-neighbour graphs are noisy, or
if $\lambda$ is too large for the local geometry, the transported point can
leave the useful region of the fitted latent space. The sensitivity and
frozen-Isomap diagnostics are included to expose this risk.

\paragraph{Privacy.}
The sampler is not differentially private, not a de-identification procedure,
and not a disclosure-control mechanism. Nearest-neighbour and overlap tests can
detect some obvious reuse behaviour, but they do not replace a threat model.

\input{tables/limitations_scope}
