\section{Implementation and Usage}
\label{sec:implementation}

The public library entry point is \texttt{DataFrameSampler}. The estimator
accepts pandas dataframes in \texttt{fit} and \texttt{transform};
\texttt{transform} returns a NumPy latent matrix; \texttt{inverse\_transform}
accepts a NumPy matrix and returns a dataframe; and \texttt{generate} returns a
synthetic dataframe. This interface is intentionally close to the scikit-learn
transformer convention, because the main implementation constraint is not to
expose several competing vectorization modes, but to make one explicit latent
construction easy to fit, inspect, persist, and reuse.

The implementation is organized around five small modules: 1) \texttt{sampler.py}
contains NCA latent learning, decoding, model persistence, and
neighbour-transport generation; 2) \texttt{neighbours.py} contains
mutual-neighbour estimation and sampling probabilities; 3) \texttt{knn.py}
contains exact, scikit-learn, and optional approximate nearest-neighbour
backends; 4) \texttt{io.py} contains CSV and Parquet dataframe I/O; and
5) \texttt{cli.py} contains the file-oriented command-line workflow. In
practice, this separation keeps optional nearest-neighbour dependencies away
from the default path while preserving a single public estimator.

The implementation therefore exposes one direct estimator and removes auxiliary
configuration, column-selection, embedding-method-selection, and bin-encoding
workflows. The CLI exposes only the NCA and neighbour-transport parameters,
decoder calibration as an opt-in diagnostic choice, rejection constraints for
generated latent candidates, and \texttt{nca\_fit\_sample\_size} as a practical
scaling parameter. Values in $(0,1]$ fit each NCA block on that fraction of
rows, and integer values fit on at most that many rows, before applying the
learned projection to the full dataframe. This is consistent with the paper's
practical-use claim: low setup cost is favoured over a large configuration
surface, while the remaining parameters correspond to explicit parts of the
method and to boundary checks that a user may need to inspect. The same fitted
object can be used from Python notebooks, persisted with the package helpers,
or called through the file-oriented command-line interface for CSV and Parquet
workflows.

For routine use, the paper names three operating points rather than asking
users to treat every parameter as a tuning axis. The fast setup
(\texttt{n\_iterations=0}, no retries, no decoder calibration) is meant for
smoke tests and quick previews. The default setup
(\texttt{n\_iterations=1}, five retries, no decoder calibration) is the
standard example-data workflow. The accurate setup
(\texttt{n\_iterations=2}, twenty retries, calibrated decoders) is slower and
intended for diagnostic runs where probability quality and stricter candidate
screening matter.
