\section{Introduction}

\paragraph{WHAT.}
This paper studies a narrow synthetic-tabular construction problem rather than
the broader problem of releasing a high-fidelity or formally private synthetic
dataset. The input is a mixed-type pandas dataframe with continuous columns,
categorical columns, booleans, and binary numeric codes; the output is another
dataframe with the same schema, intended for tests, demonstrations, notebooks,
and exploratory modelling. More precisely, the task is to generate example rows
that preserve enough schema, marginal structure, and local dependence structure
to be useful in low-stakes software workflows, while keeping the modelling
assumptions inspectable and the public interface small.

\paragraph{WHY.}
Direct reuse of source records can be infeasible even when the intended use is
modest, because governance rules, licensing restrictions, privacy concerns, or
operational separation may prevent developers from placing real rows in tests,
dashboards, training material, or notebooks. In principle, a user could turn to
a larger generative model, a disclosure-control system, or a task-specific
oversampler, however these alternatives often introduce additional dependencies,
implicit objectives, or claims that are stronger than the example-data workflow
requires. The practical failure condition is therefore not simply poor benchmark
performance; it is a mismatch between a small need, namely auditable example
generation, and tools whose evidence or configuration surface is organized
around stronger fidelity, augmentation, or release claims.

\paragraph{HOW.}
DataFrameSampler addresses this gap by using one explicit latent construction
throughout. Non-binary numeric columns are median-imputed and standardized,
while every categorical column, including booleans and numeric two-valued
columns such as $0/1$, is treated as a supervised target. For each categorical
target the method learns a Neighborhood Components Analysis block from the
remaining columns, composes these blocks with the numeric coordinates, generates
candidate latent rows by local mutual-neighbour displacement, and decodes the
result through inverse numeric scaling and random-forest categorical
classification. In practice, this gives a small estimator whose mechanism can
be inspected as fitted blocks, neighbour chains, rejection diagnostics, and
per-column decoders, rather than as a single opaque generative objective.

\paragraph{Contributions.}
The paper makes four contributions, each tied to a bounded claim. First, we
provide a direct \texttt{DataFrameSampler} estimator with \texttt{fit},
\texttt{transform}, \texttt{inverse\_transform}, and \texttt{generate}. Second,
we define an iterative per-categorical-column NCA latent construction for
mixed-type dataframes, with binary variables treated as categorical targets
rather than continuous coordinates. Third, we use a local mutual-neighbour
transport rule in the learned space as an inspectable generation heuristic,
including rejection-and-retry constraints for fitted numeric ranges and
standardized numeric outliers. Fourth, we organize the empirical section around
claim-specific evidence: distributional similarity, downstream utility,
neighbour-distance behaviour, NCA-block mechanism evidence, decoder probability
diagnostics, and extrapolative manifold consistency are interpreted only within
the conclusion each result can support.
