\section{Method}
\label{sec:method}

DataFrameSampler constructs a numerical latent space from a mixed-type
dataframe, performs local transport in that space, and decodes transported
points back into the original schema. The construction is intentionally
composite: each categorical column receives its own supervised block, while
continuous columns remain standardized coordinates. This section specifies the
latent construction, the decoder, the stochastic transport operator, and the
computational cost of the default implementation.

\subsection{Latent Construction}

Let $X^{num}$ be the median-imputed and standardized matrix of continuous
columns. Let $C_1,\ldots,C_q$ be the categorical columns under the type policy
in Section~\ref{sec:problem-setting}. Each categorical column is first encoded
as a one-hot block, and the initial context is the concatenation of
standardized continuous coordinates and all one-hot blocks.
The role of the learned blocks is to construct real-valued proxies for
categorical variables so that rows with the same categorical label are mapped
toward nearby points, making categorical variation compatible with local
geometric operations.

For each categorical column $C_j$, the sampler removes $C_j$'s own current
block from the context and fits a supervised projection
\[
  z_j = f_j(X_{-j}) \in \mathbb{R}^{k_j}
\]
with Neighborhood Components Analysis, using the labels of $C_j$ as the target.
The learned block replaces the previous block for $C_j$. This replacement is
performed for all categorical columns and repeated for a configured number of
iterations. The final latent matrix is
\[
  Z = [X^{num} \mid f_1(X_{-1}) \mid \cdots \mid f_q(X_{-q})].
\]
When the iteration count is zero, the replacement step is skipped and the
initial one-hot categorical blocks remain in the latent matrix; however, the
target-exclusion rule still applies to decoder training and diagnostics.
For larger datasets, the NCA fit can be estimated on a deterministic row
subsample and then applied to all rows. In the implementation this is controlled
by \texttt{nca\_fit\_sample\_size}: an integer gives a row cap, while a value in
$(0,1]$ gives a fraction of the fitted rows. The paper experiments use a
fraction of $0.25$ as the initial practical setting.

\subsection{Algorithms}

The following pseudocode gives the fitted construction in operational form.
It is deliberately written at the same level as the implementation, because
reproducibility is part of the contribution.

\begin{algorithm}[t]
\caption{Fit the supervised composite latent representation.}
\label{alg:fit-latent}
\begin{algorithmic}[1]
\Require Mixed-type dataframe $X$, iteration count $T$, block widths $\{k_j\}_{j=1}^q$, optional NCA fit sample size $s$
\Ensure Fitted latent matrix $Z$, categorical decoders $\{g_j\}_{j=1}^q$
\State Split columns into continuous columns and categorical columns using the binary-as-categorical policy.
\State Median-impute and standardize continuous columns to obtain $X^{num}$.
\State Fill categorical missing values with a sentinel and one-hot encode each categorical column.
\State Initialize the context $H \gets [X^{num}\mid O_1 \mid \cdots \mid O_q]$.
\For{$t = 1,\ldots,T$}
  \For{categorical column $C_j$}
    \State Remove $C_j$'s current block from $H$ to form $H_{-j}$.
    \State Select deterministic stratified row subset $I_j$ using $s$, if provided.
    \State Fit NCA projection $f_j:H_{-j}\rightarrow\mathbb{R}^{k_j}$ using labels $C_j$ on rows $I_j$.
    \State Replace $C_j$'s current block in $H$ with $f_j(H_{-j})$.
  \EndFor
\EndFor
\State Set $Z \gets H$.
\For{categorical column $C_j$}
  \State Remove $C_j$'s final block from $Z$ to form $Z_{-j}$.
  \State Fit random-forest decoder $g_j$ to predict $C_j$ from $Z_{-j}$.
\EndFor
\State \Return fitted preprocessing, projections, $Z$, and decoders $\{g_j\}_{j=1}^q$.
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
\caption{Generate synthetic rows by local displacement transfer.}
\label{alg:generate}
\begin{algorithmic}[1]
\Require Fitted latent matrix $Z$, decoders $\{g_j\}_{j=1}^q$, sample count $m$, scale $\lambda$, retry budget $R$
\Ensure Synthetic dataframe $\widetilde{X}$
\For{$i = 1,\ldots,m$}
  \State Sample anchor row $A$ from $Z$.
  \For{$r = 0,\ldots,R$}
    \State Sample a mutual neighbour $B$ of $A$.
    \State Sample a mutual neighbour $C$ of $B$.
    \State Form candidate $A' \gets A + \lambda(C-B)$.
    \If{$A'$ satisfies fitted latent range and numeric z-score constraints}
      \State \textbf{break}
    \EndIf
  \EndFor
  \State Decode continuous coordinates of $A'$ by inverse scaling.
  \State Decode each categorical column by sampling from $g_j(A'_{-j})$.
\EndFor
\State \Return decoded dataframe $\widetilde{X}$.
\end{algorithmic}
\end{algorithm}

We use three named configurations when discussing practical setup choices. The
\emph{fast} setup uses no NCA refinement iteration, no rejection retries, and
uncalibrated decoders. The \emph{default} setup uses one NCA iteration, five
rejection retries, and uncalibrated decoders. The \emph{accurate} setup uses
two NCA iterations, twenty rejection retries, and calibrated decoders. These
are not separate methods; they are speed--accuracy operating points of the
same estimator.

\subsection{Decoding}

The inverse transformation is approximate. Continuous coordinates are mapped
back through the fitted scaler, but categorical coordinates require a learned
inverse map because the NCA blocks are not invertible. For each categorical
column $C_j$, the sampler trains a \texttt{RandomForestClassifier} to predict
the labels of $C_j$ from the final latent context with $C_j$'s own block
removed. Thus the decoder follows the same exclusion rule as the encoder:
the category being reconstructed is never provided as an input feature.

At reconstruction time, the classifier returns a probability vector over
categories observed during fitting. The decoder can either sample from that
distribution or select the maximum-probability category. Probability
calibration is optional; when enabled and class counts permit it, the forest is
wrapped with scikit-learn's \texttt{CalibratedClassifierCV}. Calibration quality
is still measured empirically, because calibrated model probabilities are not
equivalent to a proof of conditional distribution correctness.

\subsection{Stochastic Transport}

Generation is a stochastic operator over the fitted latent matrix. For an
anchor $A$, a mutual neighbour $B$, and a mutual neighbour $C$ of $B$, the
candidate point is
\[
  T(A;B,C)=A+\lambda(C-B).
\]
The vector $C-B$ is a local displacement observed near $B$ and transferred to
the neighbourhood of $A$. This construction differs from interpolation:
for $\lambda>0$, $T(A;B,C)$ need not lie on the segment between $A$ and a
neighbour of $A$, and it can lie outside the convex hull of the training latent
rows. The operator is nevertheless only a heuristic. It is meaningful when the
learned latent space is locally regular enough for nearby displacements to be
reused, and it can fail when this local-linearity assumption is violated.

The same formula is also the method's main interpretability interface. If
$D=T(A;B,C)$, then the generated point can be read as the observed-displacement
analogy
\[
  D : A \approx C : B,
\]
or, in words, ``modify $A$ in the same fitted-latent direction by which $C$
differs from $B$.'' This is an analogy over measured rows and fitted latent
coordinates, not a semantic or causal statement about the underlying domain.
Its practical value is that every generated row has a trace: the anchor $A$,
the observed pair $(B,C)$, the displacement $C-B$, the accepted or rejected
candidate, and the decoder probabilities used for categorical reconstruction.

\subsection{Computational Complexity}

Let $n$ be the number of fitted rows, $s$ the number of rows used to fit each
NCA block after optional subsampling, $p$ the number of continuous columns,
$q$ the number of categorical columns, $h$ the total one-hot width, $k_j$ the
latent width for categorical column $j$, and $T$ the number of NCA refinement
iterations. Median imputation, one-hot encoding, and standardization are
linear in the number of observed dataframe entries. Exact NCA is the dominant
default cost: for each categorical column and iteration, scikit-learn NCA
optimizes a supervised neighbourhood objective whose distance computations are
quadratic in $s$ and linear in the current context width. A coarse upper bound
for the NCA phase is therefore
\[
  O\!\left(T q s^2 (p + h + \sum_j k_j)\right),
\]
up to optimizer constants and effective context widths. With no NCA subsampling,
$s=n$; with a fractional setting such as \texttt{nca\_fit\_sample\_size=0.25},
$s\approx 0.25n$, while the learned projection is still evaluated on all $n$
rows.

Decoder fitting trains one random forest per categorical column and is
parallelized across trees by default. Exact nearest-neighbour construction in
the final latent space is quadratic in $n$ for brute-force search, while
optional approximate backends can reduce practical runtime at the cost of an
additional approximation choice. Once fitted neighbours are available,
generation is inexpensive: each sample requires a small number of neighbour
lookups, one latent displacement, constraint checks over latent coordinates,
and one decoder pass per categorical column. The method is therefore designed
for small and medium notebook-scale data by default, with approximate
neighbour search left as an optional scaling path.

The main memory cost is storing the fitted context matrices and learned latent
blocks, which is linear in $n$ times the active context width, plus the
per-tree storage of the categorical random forests. Exact NCA also materializes
neighbourhood computations whose practical footprint grows quickly with $n$.
For this reason, the implementation is aimed at notebook-scale and
medium-sized tabular workflows. Very large tables should be handled with row
caps, representative subsampling, narrower categorical preprocessing, or
approximate neighbour search rather than by treating the default exact path as
a large-scale generative learner.
