\section{Related Work}
\label{sec:related-work}

Synthetic tabular data generation includes several distinct modelling
traditions. Copula and probabilistic models provide relatively classical
approaches to mixed marginal and dependence modelling, while systems such as
the Synthetic Data Vault expose practical interfaces over multiple synthetic
data model families \citep{patki2016sdv}. Deep generative tabular models, such
as CTGAN, use adversarial training and conditional sampling to model more
complex tabular distributions \citep{xu2019ctgan}. These methods are important
reference points, but they are usually optimised for broader fidelity or
release-oriented objectives than the low-setup, inspectable example-data
workflow considered here.

\begin{table}[t]
\centering
\caption{Positioning against common tabular-generation families. The entries
are schematic rather than benchmark claims; their role is to locate the method
by modelling objective and inspectability. Takeaway: DataFrameSampler trades a
global generative objective for explicit local geometry and a small
mixed-type workflow.}
\label{tab:method-positioning}
\small
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{@{}llllll@{}}
\toprule
Method & Objective & Mixed-type handling & Local geometry & Interpretability & Setup cost \\
\midrule
CTGAN & Adversarial fidelity & Learned & Implicit & Low & High \\
Copula & Likelihood/dependence & Structured marginals & None & Medium & Low \\
SMOTE-style & Interpolation & Numeric or encoded & Linear segments & High & Low \\
DataFrameSampler & Constructive, no global loss & Explicit type policy & Learned neighbour transport & High & Low \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

Oversampling methods provide a second reference point because they make the
geometry of synthetic example construction explicit. SMOTE generates minority
examples by interpolation along line segments between neighbouring observations
\citep{chawla2002smote}. This rule is efficient and inspectable, but it is
confined to interpolation in the chosen feature space. Manifold-based
oversampling methods address this limitation by treating the data distribution
as supported by a lower-dimensional structure, including early manifold-based
sampling and later manifold-conformance estimation
\citep{nguyen2005manifold,bellinger2018manifold}.

Recent manifold augmentation methods make the local-geometry assumption more
explicit. First-order manifold augmentation samples from estimated tangent
planes, while curvature-enhanced sampling incorporates second-order
information for regression augmentation
\citep{kaufman2024foma,kaufman2025cems}. DataFrameSampler is adjacent to this
family but uses a different constructive choice: rather than estimating a
tangent plane or curvature tensor, it transfers observed neighbour
displacements in a supervised latent space. The result is less formal than
explicit differential-geometry methods, but it is easier to express as fitted
blocks, neighbour chains, and decoder distributions.

Metric learning supplies the other main connection. Neighborhood Components
Analysis learns a supervised transformation that favours neighbours with the
same label, and scikit-learn provides the implementation used here
\citep{pedregosa2011scikit}. DataFrameSampler applies this idea columnwise:
each categorical column defines a supervised prediction problem, and the
resulting blocks are concatenated into a composite latent space. This differs
from a joint metric-learning objective, because there is no single loss over
all columns and all generated rows.

The manifold validation protocol draws on Isomap and out-of-sample manifold
extension. Isomap constructs a neighbourhood graph, computes graph shortest
paths, and embeds the resulting geodesic distances by multidimensional scaling
\citep{tenenbaum2000isomap}. Out-of-sample extensions allow new points to be
inserted into a fixed spectral geometry without refitting the original model
\citep{bengio2003outofsample}. In this paper, that machinery is used only as a
diagnostic: generated latent points are tested for convex-hull extrapolation
and for low insertion stress relative to held-out real points.

Finally, privacy-preserving synthetic data addresses a stronger problem than
the one considered here. Formal release mechanisms require explicit threat
models, privacy definitions, and disclosure-risk evaluation. DataFrameSampler
does not provide these guarantees, and row novelty under a learned decoder
should not be interpreted as privacy protection.
